ArKanjo: a tool for detecting function-level Code Duplication in the Linux Kernel

### 2025-06-18 paper/paper.tex
Reviewing the results and conclusion sections, and other minor adjustments

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Luan Ícaro Pinto Arcanjo}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil\\
Email: luanicaro@usp.br}
\and
\IEEEauthorblockN{David Tadokoro}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil\\
Email: davidbtadokoro@usp.br}
\and
\IEEEauthorblockN{Paulo Meirelles}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil\\
Email: paulormm@ime.usp.br}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{diagrama_mestrado}
\caption{Architecture diagram demonstrating the relationship between the tool components}
\label{fig:diagram}
\end{figure}

The tool operates in two main parts: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} performs heavy computations to find code 
duplications across a codebase and produces artifacts to store duplication-related information 
in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
by the Preprocessor to execute the tool functionalities as requested. This design allows the 
tool to perform the heavy and time-consuming steps only once per codebase, enabling fast performance 
for multiple queries on the same codebase.

The Preprocessor contains two main components: the Function Breaker and the Code Duplication Finder. 
The workflow of the Preprocessor is as follows: the Function Parser receives the input codebase,
extracts the functions along with metadata, and creates a temporary codebase where each extracted 
function becomes a new code file. The Code Duplication Finder then iterates over every pair of files 
in the temporary codebase, checks whether they are code clones, and, if so, stores the result in 
the Code Duplication Database. This database is a text file that records each code duplication as a triple 
\textbf{<function1, function2, similarity>}, where \textbf{function1} and \textbf{function2} are the 
duplicated functions, and \textbf{similarity} is the score returned by the duplication detection method 
used in the Code Duplication Finder. Figure~\ref{fig:diagram} illustrates the tool architecture.

The Query Responder uses the temporary codebase and the Code Duplication Database to extract 
information about duplicated functions based on the user's request. If the user executes the 
Query Responder without previously executing the Preprocessor, the Query Responder automatically 
calls the Preprocessor to generate the required artifacts.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{explorer_example}
\caption{Example of the \textit{Duplication Explorer} functionality of the ArKanjo tool.}
\label{fig:explorer_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{function_example}
\caption{Example of the \textit{Function Information} functionality of the ArKanjo tool.}
\label{fig:function_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{relatory_example}
\caption{Example of the \textit{Duplication Report} functionality of the ArKanjo tool.}
\label{fig:relatory_ex}
\end{figure}

@@

@@

\centering

\renewcommand{\arraystretch}{1.3}
\begin{tabular}{ | m{13mm} | c | c | c | c | m{6mm} | m{6mm} | }

@@

\hline
\end{tabular}
\caption{Recall of the ArKanjo tool in the BigCloneBench dataset varying minimum similarity threshould.}
\label{tab:bigclone}
\end{table}

@@

\centering

\begin{tabular}{ | m{15mm} | c | c | c | c | m{6mm} | m{10mm} | }

@@

\end{tabular}
\caption{Results of the ArKanjo tool in the empirical analyses in the AMD Display driver}
\label{tab:emp}

@@

Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCLoneBench dataset
and table \ref{tab:emp} shows the results obtained in the empricial analysis.
The BigCloneBench results and our empirical methods indicate that our tool performs
well in detecting Type-1 and Type-2 code clone duplications and can thus reveal propagated 
issues like copied buggy logic, but it struggles with more
complex types. A similarity threshold between 80\% and 100\% yields favorable results
in both methods.

However, there is a discrepancy in the detection of non-duplication pairs between
the two methods. In the BigCloneBench evaluation, we found no false positives (negative
samples inferred as duplications), while the empirical analysis revealed a higher percentage
of false positives. We propose two potential explanations for this discrepancy.

The first reason may be the limitations and known issues with BigCloneBench, as
highlighted by Krinke and Ragkhitwetsagul \cite{bigfail}. The
second reason could relate to the nature of the AMD Display driver, where each code artifact
shares semantic meaning. In contrast, BigCloneBench comprises self-contained code
artifacts without shared semantics. Given that our tool relies on a text-based code clone
detection approach, it may naturally perform worse on the AMD Display driver than on the
BigCloneBench dataset.

@@

Using the ArKanjo tool, the author and 12 student groups identified and refactored duplicated 
code in the AMD Display driver and the Industrial I/O (IIO) subsystems, submitting their work 
as patches to the community maintainers. This effort resulted in 13 total patch submissions: 
one from the author to the AMD Display driver and 12 from the student groups, with 11 targeting 
the IIO subsystem and one for the AMD Display driver. The author's patch, which removed over 
500 lines of duplicated code, was successfully accepted and merged into the kernel after a 
lengthy review process.

The student efforts showed that newcomers could use the tool to make effective contributions. 
Of the 12 student patches, five were accepted, four were refused, and three required follow-up 
fixes based on maintainer feedback. To remove the duplicated code, seven student groups used the 
Parameterize Method and two used the Extract Method \cite{refactorbook}, which are simple refactoring methods. 
This result corroborates that ArKanjo finds actionable duplications and that newcomers can 
successfully contribute patches to the kernel.

While this research included a successfully merged patch by the author and students,
a notable portion of student efforts encountered
significant hurdles. These difficulties often stemmed from maintainer feedback where
proposed changes, though reducing duplication, were rejected or required substantial
rework due to concerns about code readability, added abstraction, or the specific context
of the code. Technical complexities, such as C macros in configuration files, and lengthy
patch review times also contributed to the challenges. 
This highlights that while ArKanjo effectively identifies duplications, a successful 
contribution involves a trade-off between code deduplication and other factors like the 
specific code's context and maintainer priorities.

@@

In conclusion, this paper presented ArKanjo, a novel tool designed to effectively detect 
and facilitate the refactoring of function-level code duplication within the Linux kernel. 
Evaluations demonstrated that ArKanjo is highly effective at identifying Type-1 and Type-2 
clones, which are often indicative of propagated issues like copied buggy logic. The practical 
value of the tool was confirmed through an ethnographic study where the author and student 
groups used ArKanjo to identify actionable duplications. This effort led to 13 patch 
submissions, including one from the author that removed over 500 lines of code and was merged 
into the kernel, alongside three successful student contributions. While some refactoring 
efforts were declined due to maintainer priorities, ArKanjo successfully lowers the barrier 
for new contributors to make meaningful improvements. The tool provides immediate value for 
kernel maintenance and serves as a model for clone detection in other large-scale projects.

@@

This study was financed, in part, by  the University of
São Paulo/USP (Proc. 22.1.9345.1.2), the São Paulo Research Foundation/FAPESP
and the São Paulo State Data Analysis System Foundation/SEADE (Proc. 2023/18026-8), Brazil.

>>>>

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Luan Ícaro Pinto Arcanjo}
\IEEEauthorblockA{
Universidade de São Paulo, Brazil\\
luanicaro@usp.br}
\and
\IEEEauthorblockN{David Tadokoro}
\IEEEauthorblockA{
Universidade de São Paulo, Brazil\\
davidbtadokoro@usp.br}
\and
\IEEEauthorblockN{Paulo Meirelles}
\IEEEauthorblockA{
Universidade de São Paulo, Brazil\\
paulormm@ime.usp.br}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{diagrama_mestrado}
\caption{Architecture diagram with the tool components}
\label{fig:diagram}
\end{figure}

The tool operates in two main parts, as illustrates in Figure~\ref{fig:diagram}: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} performs heavy computations to find code 
duplications across a codebase and produces artifacts to store duplication-related information 
in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
by the Preprocessor to execute the tool functionalities as requested. This design allows the 
tool to perform the heavy and time-consuming steps only once per codebase, enabling fast performance 
for multiple queries on the same codebase.

The Preprocessor contains two main components: the Function Breaker and the Code Duplication Finder. 
The workflow of the Preprocessor is as follows: the Function Parser receives the input codebase,
extracts the functions along with metadata, and creates a temporary codebase where each extracted 
function becomes a new code file. The Code Duplication Finder then iterates over every pair of files 
in the temporary codebase, checks whether they are code clones, and, if so, stores the result in 
the Code Duplication Database. This database is a text file that records each code duplication as a triple 
\textbf{<function1, function2, similarity>}, where \textbf{function1} and \textbf{function2} are the 
duplicated functions, and \textbf{similarity} is the score returned by the duplication detection method 
used in the Code Duplication Finder.

The Query Responder uses the temporary codebase and the Code Duplication Database to extract 
information about duplicated functions based on the user's request. If the user executes it
without previously executing the Preprocessor, the Query Responder automatically 
calls the Preprocessor to generate the required artifacts.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{explorer_example}
\caption{Example of the \textit{Duplication Explorer} functionality of the ArKanjo tool.}
\label{fig:explorer_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{function_example}
\caption{Example of the \textit{Function Information} functionality of the ArKanjo tool.}
\label{fig:function_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{relatory_example}
\caption{Example of the \textit{Duplication Report} functionality of the ArKanjo tool.}
\label{fig:relatory_ex}
\end{figure}

@@

In this section, we assess the ArKanjo tool effectiveness in detecting code duplications through benchmark comparison and empirical analysis. We also describe findings from an ethnographic study involving real contributions to the Linux kernel, highlighting the tool practical utility and limitations in developer workflows.

@@

\centering
\caption{Recall of the ArKanjo tool in the BigCloneBench dataset.}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{ | m{13mm} | c | c | c | c | m{6mm} | m{6mm} | }

@@

\hline
\end{tabular}

\label{tab:bigclone}
\end{table}

@@

\centering
\caption{Results of the ArKanjo tool analyzing the AMD Display driver.}
\begin{tabular}{ | m{15mm} | c | c | c | c | m{6mm} | m{10mm} | }

@@

\end{tabular}

\label{tab:emp}

@@

Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCloneBench dataset,
and Table \ref{tab:emp} presents the results from the empirical analysis.
The outcomes from both the BigCloneBench evaluation and our empirical method indicate that
our tool performs well in detecting Type-1 and Type-2 code clone duplications,
which are sufficient to reveal propagated issues like copied buggy logic.
However, it struggles with more complex clone types. A similarity threshold between 80\% and 100\% 
yields favorable results in both methods.

Nevertheless, a discrepancy emerges when detecting non-duplicate pairs between
the two methods. In the BigCloneBench evaluation, we found no false positives (i.e., negative
samples inferred as duplications), whereas the empirical analysis revealed a higher percentage
of false positives. We propose two potential explanations for this discrepancy.

The first explanation concerns the limitations and known issues with BigCloneBench, as
highlighted by Krinke and Ragkhitwetsagul \cite{bigfail}. The
second explanation relates to the nature of the AMD Display driver, where code artifacts
often share semantic meaning. In contrast, BigCloneBench comprises self-contained code
artifacts with minimal or no shared semantics. Since our tool relies on a text-based code clone
detection approach, it may naturally perform worse on the AMD Display driver than on the
BigCloneBench dataset.

@@

Using the ArKanjo tool, the main author and 12 student groups identified and refactored duplicated 
code in the AMD Display driver and the Industrial I/O (IIO) subsystems, submitting their work 
as patches to the community maintainers. This effort resulted in 13 patch submissions: 
One was from the main author to the AMD Display driver, and 12 were from the student groups, with 11 targeting 
the IIO subsystem and one aimed at the AMD Display driver. The author's patch was successfully accepted and merged into the kernel after a 
lengthy review process, removing over 
500 lines of duplicated code.

The student efforts demonstrated that newcomers could use the tool to contribute effectively. 
Of the 12 student patches, five were accepted, four were rejected, and three required follow-up 
fixes based on maintainer feedback. To remove the duplicated code, seven student groups used the 
\textit{Parameterize Method}, and two used the \textit{Extract Method} \cite{refactorbook}, both of which are 
straightforward refactoring strategies. These results support the conclusion that ArKanjo identifies 
actionable duplications and that newcomers can successfully contribute patches to the kernel.

While this research includes successfully merged patches by the main author and student groups,
a significant portion of the student contributions encountered notable challenges. These difficulties 
often stemmed from maintainer feedback indicating that, although the proposed changes reduced duplication, 
they were rejected or required substantial revision due to concerns about code readability, increased abstraction, 
or the specific context of the code. Additional challenges included technical complexities, such as C macros in 
configuration files and extended patch review timelines.

These findings highlight that while ArKanjo effectively identifies code duplications, a successful 
contribution also requires navigating trade-offs between deduplication and other factors, such as 
contextual appropriateness and maintainer expectations.

@@

This paper presented ArKanjo, a novel tool designed to detect effectively 
and facilitate the refactoring of function-level code duplication within the Linux kernel. 
Evaluations demonstrated that ArKanjo is highly effective at identifying Type-1 and Type-2 
clones, which are often indicative of propagated issues such as copied buggy logic. We validated the practical 
value of the tool through an ethnographic study in which the main author and student 
groups used ArKanjo to identify actionable duplications. This effort led to 13 patch 
submissions, including one from the main author that removed over 500 lines of code and merged 
into the kernel, alongside five successful student contributions at this moment. While some refactoring 
efforts were declined due to maintainer priorities, ArKanjo successfully lowered the barrier 
for new contributors to make meaningful improvements. The tool provides immediate value for 
kernel maintenance and serves as a model for clone detection in other large-scale projects.

@@

This study was financed, in part, by CAPES (Finance Code 001), the University of São Paulo – USP (Proc. 22.1.9345.1.2), the São Paulo Research Foundation – FAPESP and the São Paulo State Data Analysis System Foundation – SEADE (Proc. 2023/18026-8), Brazil.
